---
title: "Gideon_Reproducible Research Course Project#1"
author: "Gideon Obeisun"
date: "August 2, 2017"
output:
  pdf_document: default
  html_document: default
---

### Loading the activity monitoring dataset into R dataframe

```{r Monitoring_Activity_Data dataframe, echo=TRUE}

	setwd("C:/Users/Gideon87/Documents/R/Cousera Reproducible Research Course")

Activity_Data<- read.csv("activity.csv")


```

## Question 1 - What is mean total nubers of steps taken per day?

### Calculate the total no. of steps per day.
```{r Total_steps/day,echo=TRUE}
Activity_Data$date<-as.Date(Activity_Data$date,format="%Y-%m-%d")
Activity_Data<-data.frame(Date=Activity_Data$date,Dayof_Week=tolower(weekdays(Activity_Data$
               date)),Steps=Activity_Data$steps,Interval=Activity_Data$interval)
Typeof_Day<-ifelse(Activity_Data$Dayof_Week=="saturday"|Activity_Data$Dayof_Week=="sunday",
                   "weekend","weekday")
Activity_Data<-cbind(Activity_Data,Typeof_Day)
Activity_Data2 <- data.frame(Date =Activity_Data$Date,
                       Dayof_Week =Activity_Data$Dayof_Week,
                       Typeof_Day =Activity_Data$Typeof_Day,
                       Interval =Activity_Data$Interval,
                       Steps =Activity_Data$Steps)
Steps_perDay<-aggregate(Activity_Data2$Steps,by=
                      list(Activity_Data2$Date),FUN=sum,na.rm=TRUE)

names(Steps_perDay)<-c("Date","Total")
                                                                                  
                                                              
```
### The histogram of the total number of steps taken each day.
```{r Histogram of Total_Steps/Day,echo=TRUE}

hist(Steps_perDay$Total,main="Total Number of Steps Taken Per Day"
     ,xlab="Total_Steps/Day",breaks=seq(from=0,to=30000,by=2000))
#save the graph as a png file.
dev.copy(png, file="PA1_plot1.png",height=500,width=500)

#shutdown the graphic device.
dev.off()
```
### Calculating and reporting the mean and median of the total numbers of steps taken
### per day.
```{r Mean/Median of Total_Steps/Day,echo=TRUE}

print(Mean_Steps_Per_Day<-mean(Steps_perDay$Total, na.rm =TRUE))
print(Median_Steps_Per_Day<-median(Steps_perDay$Total,na.rm = TRUE))
```
## Question 2 - What is the average daily activity pattern?

### Time series plot of 5-minute interval/avg.no.of steps taken across all days.
```{r Time-Series Plot,echo=TRUE}

Mean_Steps_AllDay<-aggregate(Activity_Data2$Steps,by=list(Activity_Data2$Interval),FUN=mean,
                            na.rm=TRUE)
names(Mean_Steps_AllDay)<-c("Interval","Mean_Step")
with(Mean_Steps_AllDay,{
  plot(Mean_Step~Interval,type="l",main="5-Minute Intervals and Average Numbers of Steps
       taken across all Days",xlab="5-minute interval(mins)",ylab="average no.of steps
       taken across all days")

}
     )
#save the graph as a png file.
dev.copy(png, file="PA1_plot2.png",height=500,width=500)

#shutdown the graphic device.
dev.off()
```
### The 5-minute interval with maximum number of steps across all days.
```{r 5-minute interval with max.no.of steps,echo=TRUE}
Max_Step<-max(Mean_Steps_AllDay$Mean_Step)

Max_Interval<-subset(Mean_Steps_AllDay,Mean_Steps_AllDay$Mean_Step==Max_Step,
                     select=Interval)
```
### Calculating and reporting the numberof missing values in the dataset.
```{r Missing Values,echo=TRUE}
NA_Activity<-subset(Activity_Data2,is.na(Steps),select=c(Steps,Date))
NA_Activity_Count<-nrow(NA_Activity)
```
### Filling in the missing values in the dataset.
```{r Filling Missing Values with Zero,echo=TRUE}
NA_Activity_Position <- which(is.na(Activity_Data2$Steps))
Mean_Values <- rep(mean(Activity_Data2$Steps, na.rm=TRUE), times=length(NA_Activity_Position))
Activity_Data2[NA_Activity_Position, "Steps"] <- Mean_Values
head(Activity_Data2)

```
### Histogram of the total number of steps taken each day with missing values of steps Changed
### to zeroes.
```{r Histogram with Zeroes for missing values.}
Non_Missing_Steps_Per_Date <- aggregate(Activity_Data2$Steps, by=list(Activity_Data2$Date), FUN=sum)
names(Non_Missing_Steps_Per_Date) <- c("Date", "Total_Steps")

hist(Non_Missing_Steps_Per_Date$Total_Steps,breaks=seq(from=0, to=30000, by=2000),xlab="Total       number of steps",ylim=c(0,40), 
     main="Histogram of the total number of steps taken each day with NA replaced by mean          value")
#save the graph as a png file.
dev.copy(png, file="PA1_plot3.png",height=480,width=480)

#shutdown the graphic device.
dev.off()
```
### It look like imputting values for the missing data created a graph different from the 
### previous histogram depending on the variances of the substituted values from the rest of the step values provided. 

## Question 3 - Are there differences in activity patterns between weekdays and weekends?

### A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r Panel Plot,echo=TRUE}
library(lattice)
Mean_Values2 <- aggregate(Activity_Data2$Steps,by=list(Activity_Data2$Typeof_Day, 
                               Activity_Data2$Dayof_Week, Activity_Data2$Interval), mean)

names(Mean_Values2) <- c("Typeof_Day", "Dayof_Week", "Interval", "Mean_Value")

xyplot(Mean_Value ~ Interval | Typeof_Day,Mean_Values2, 
       type = "l", 
       lwd = 0.5, 
       xlab = "Interval", 
       ylab = "Number of Steps", 
       layout = c(1, 2))
#save the graph as a png file.
dev.copy(png, file="PA1_plot4.png",height=480,width=480)

#shutdown the graphic device.
dev.off()
```




